---
title: "02-Subsetting"
author: "Stas Sajin"
date: "`r Sys.Date()"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Quiz

###1 Fix each of the following common data frame subsetting errors:

```{r}
#mtcars[mtcars$cyl = 4, ]
mtcars[mtcars$cyl == 4, ]

#mtcars[-1:4, ]
mtcars[1:4, ] #don't mix negative and positive subsetting

#mtcars[-1:4, ]
mtcars[mtcars$cyl <= 5,] #needs comma for logical subsetting

#mtcars[mtcars$cyl == 4 | 6, ]
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
```


###2 Why does x <- 1:5; `x[NA]` yield five missing values? (Hint: why is it different from `x[NA_real_]`?)

the `x[]` returns the original vector. Since the index is NA, it returns NA for each of the values in the original vector

the `x[NA_real_]` returns a one length vector, since the NA here is a typeof numeric. Hence, the NA subsetting returns NA for only one value in the vector.


###2.3 When should you use drop = FALSE?
It is used when you want to keep the structure of the output the same as the input. Hadley seems to strongly recommend that drop=FALSE be used for factor/array/dataframe subsetting. 

###3 If x is a matrix, what does x[] <- 0 do? How is it different to x <- 0?
```{r}
#create a 3x3 matrix
x<-matrix(data=c(1:9),nrow = 3,ncol=3)
x
#this should replace everything with 0
x[] <- 0 
x

#this replaces the matrix stored in x with a 0.
x <- 0
x
```

###4 How can you use a named vector to relabel categorical variables?
I'm not sure what is being asked here.


## Exercises

### 1--Same as quiz question 1
### 2--Same as quiz question 2
### 3 What does `upper.tri()` return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?


```{r}
#upper.tri() returns  a matrix of logicals the same size as the given matrix in the lower or upper triangle. Using the upper.tri() for subsetting will return the values for the upper triangle. We can specify the rule for whether it should also subset the diagonal or not.
x <- outer(1:5, 1:5, FUN = "*")
x
x[upper.tri(x)]
```

### 4 Why does mtcars[1:20] return an error? How does it differ from the similar mtcars[1:20, ]?

It returns an error because the column is not specified. The second expression specifies that all colums should be returned.

### 5 Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).
```{r}
x<-matrix(1:16,4)

diagonal<- function(matrix){
    x<-c() #create empty vector
    for (i in 1:nrow(matrix)) {
        x <- c(x, matrix[i,i])
    }
    return(x)
}

x
diagonal(x)
diag(x)
```

###6 What does `df[is.na(df)] <- 0` do? How does it work?
The function replaces the NAs in the dataframe with 0. First, `is.na(df)` identifies the indixes for the cells that have NA. It returns TRUE for every cell in the df that has NA and FALSE otherwise. The [], then simply subsets the TRUE values. the `<-0` replaces the subsetted values with 0.


### 7. Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Extract the R squared from the model summary (summary(mod))

```{r}
mod <- lm(mpg ~ wt, data = mtcars)

#extract df.residual
mod$df.residual

#extract r-squared value
summary(mod)$r.squared
```


### 8. How would you randomly permute the columns of a data frame? (This is an important technique in random forests.) Can you simultaneously permute the rows and columns in one step?

```{r}
#let's create a dataframe first
df <- data.frame(matrix(1:100,10))
df

#let's randomply shuffle the rows
df[sample(nrow(df)),]

#now both rows and columns
df[sample(nrow(df)),sample(ncol(df))]
```

### 9. How would you select a random sample of m rows from a data frame? What if the sample had to be contiguous (i.e., with an initial row, a final row, and every row in between)?

```{r}
#select 6 rows
m<-6
df[sample(nrow(df),m),]

#sample the first position of a row
firstPosition<-sample(nrow(df)-m+1,1)
lastPosition<-firstPosition-1+m
df[firstPosition:lastPosition,]
```


### 10. How could you put the columns in a data frame in alphabetical order?
I would use something along the lines: `names(df) <- order(names(df))`