---
title: "09-Functional Programming"
author: "Stas Sajin"
date: "`r Sys.Date()"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pryr)
```



### 1. Given a function, like "mean", match.fun() lets you find a function. Given a function, can you find its name? Why doesn't that make sense in R?

The short answer is no. In R functions are objects in their own right and are not automatically bound to a name. COnsider the case of anonymous functions that have no name.

### 2. Use lapply() and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the mtcars dataset.

```{r}
lapply(mtcars, function(x) sd(x)/mean(x))
```

### 3. Use integrate() and an anonymous function to find the area under the curve for the following functions. Use Wolfram Alpha to check your answers.

1. y = x ^ 2 - x, x in [0, 10]
2. y = sin(x) + cos(x), x in [-pi, pi]
3. y = exp(x) / x, x in [10, 20]

```{r}
#for the first case
integrate(function(x) x ^ 2 - x, 0, 10)
#for the second case
integrate(function(x) sin(x) + cos(x), -pi, pi)
#for the third case
integrate(function(x) exp(x) / x, 10, 20)
```

### 4. A good rule of thumb is that an anonymous function should fit on one line and shouldn't need to use {}. Review your code. Where could you have used an anonymous function instead of a named function? Where should you have used a named function instead of an anonymous function?

It sould be beneficial to create a named function for standard error. Other than that, we don't really have a need to use anonymous functions for integration because we could just replace the x values with the lower and upper bound we are integrating on. 


### 5. Why are functions created by other functions called closures?
Closures enclose the environment of a parent function and access all of its variables. 

### 6. What does the following statistical function do? What would be a better name for it? (The existing name is a bit of a hint.)

The formula `(x^lambda-1)/lambda` is for Box-Cox transformation. This function will return a function that that either returns the log or computes the box-cox transform. 
```{r}
bc <- function(lambda) {
  if (lambda == 0) {
    function(x) log(x)
  } else {
    function(x) (x ^ lambda - 1) / lambda
  }
}
```

### 7. What does approxfun() do? What does it return?
According to the documentation: "Returns a list of points which linearly interpolate given data points, or a function performing the linear (or constant) interpolation.""

### 8. What does ecdf() do? What does it return?
Documentation: "Computes an empirical cumulative distribution function, with several methods for plotting, printing and computing with such an "ecdf" object."

### 9. Create a function that creates functions that compute the ith central moment of a numeric vector. You can test it by running the following code:
```{r}
#here is the function
moment<-function(momentOrder) {
    function (x) (sum((x-mean(x))^momentOrder)/length(x))
}

m1 <- moment(1)
m2 <- moment(2)

unenclose(m1)
unenclose(m2)

x <- runif(100)
stopifnot(all.equal(m1(x), 0))
stopifnot(all.equal(m2(x), var(x) * 99 / 100))
#all tests have passed
```

### 10. Create a function pick() that takes an index, i, as an argument and returns a function with an argument x that subsets x with i.

```{r}
#here is the solution:
pick <- function(i) { function(x) x[[i]] }

lapply(mtcars, pick(5))
# should do the same as this
lapply(mtcars, function(x) x[[5]])
```


### 11. Implement a summary function that works like base::summary(), but uses a list of functions. Modify the function so it returns a closure, making it possible to use it as a function factory.

### 12. Which of the following commands is equivalent to with(x, f(z))?
1. x$f(x$z).
2. f(x$z).
3. x$f(z).
4. f(z).
5. It depends.

### 13. Instead of creating individual functions (e.g., midpoint(), trapezoid(), simpson(), etc.), we could store them in a list. If we did that, how would that change the code? Can you create the list of functions from a list of coefficients for the Newton-Cotes formulae?

### 14. The trade-off between integration rules is that more complex rules are slower to compute, but need fewer pieces. For sin() in the range [0, ??], determine the number of pieces needed so that each rule will be equally accurate. Illustrate your results with a graph. How do they change for different functions? sin(1 / x^2) is particularly challenging.